---
title: "ALE_Experiment"
author: "Kenzie Cooke"
date: "2024-06-16"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

## Alkalinity and Linear Extension Experient

```{r libraries}
library(ggplot2)
library(dplyr)
library(nlstools)
library(seacarb)
library(emmeans)

```

### Import Data
*bw data normalized to initial mass and not including density corrections

```{r import, echo=FALSE}
#import growth data 
bw <- read.csv('data/ALEbw6.16.csv', stringsAsFactors = TRUE)
le <- read.csv('data/ALE_le6.16.csv', stringsAsFactors = TRUE)
```

### Mixed linear models for total calcification by TA
1. BW from t0 to t3, growth ~ ta, 1|geno, 1|tank

```{r}
#attempting to fix "error in initializerPtr(): function 'chlomod_factor_ldetA' not provided by package 'Matrix'

#tools::package_dependencies("Matrix", which = "LinkingTo", reverse = TRUE)[[1L]]

#utils::install.packages("lme4", type = "source")

#load packages
library(lmerTest)

#BW mixed linear model 1
LM_BWTA1 <- lmerTest::lmer(gr_t0t3 ~ ta + (1|geno) + (1|tank), data = bw)
summary(LM_BWTA1)
anova(LM_BWTA1)
ranova(LM_BWTA1)
step(LM_BWTA1)

LM_BWTA1b <- lm(gr_t0t3 ~ ta, data = bw)
summary(LM_BWTA1b)
anova(LM_BWTA1b)

```
2. BW from t1 to t3

```{r}
#BW mixed linear model 2
LM_BWTA2 <- lmer(gr_t1t3 ~ ta + (1|geno) + (1|tank), data = bw)
summary(LM_BWTA2)
anova(LM_BWTA2)
```
#### By Treatment 
1. BW from t0 to t3
```{r}
#BW mixed linear model by treatment 
LM_BWtreat1 <- lmer(gr_t0t3 ~ treatment + (1|geno) + (1|tank), data = bw)
summary(LM_BWtreat1)
anova(LM_BWtreat1)
```
2. BW from t1 to t3
```{r}
#BW mixed linear model 1
LM_BWtreat2 <- lmer(gr_t1t3 ~ treatment + (1|geno) + (1|tank), data = bw)
summary(LM_BWtreat2)
anova(LM_BWtreat2)
```


### Mixed linear models for linear extension by TA
1. LE (leios scanning)

```{r}
#mixed linear model 1
LM_LETA1 <- lmer(gr_leios ~ ta + (1|geno) + (1|tank), data = le)
summary(LM_LETA1)
anova(LM_LETA1)
ranova(LM_LETA1)
```

2. LE from t0 to t3 using ImageJ

```{r}
#mixed linear model 2
LM_LETA2 <- lmer(gr_t0t3 ~ ta + (1|geno) + (1|tank), data = le)
summary(LM_LETA2)
anova(LM_LETA2)
ranova(LM_LETA2)

LM_LETA2 <- lmer(gr_t0t3 ~ ta * geno + (1|tank), data = le)
summary(LM_LETA2)
anova(LM_LETA2)
step(LM_LETA2)
```

3. LE from t1 to t3 using Image J

```{r}
#mixed linear model 3
LM_LETA3 <- lmer(gr_t1t3 ~ ta + (1|geno) + (1|tank), data = le)
summary(LM_LETA3)
anova(LM_LETA3)
```
#### By Treatment 
1. LE by leios
```{r}
#mixed linear model 1
LM_LEtreat1 <- lmer(gr_leios ~ treatment + (1|geno) + (1|tank), data = le)
summary(LM_LEtreat1)
anova(LM_LEtreat1)
```

2. LE t0 to t3 Image J 
```{r}
#mixed linear model 2
LM_LEtreat2 <- lmer(gr_t0t3 ~ treatment + (1|geno) + (1|tank), data = le)
summary(LM_LEtreat2)
anova(LM_LEtreat2)
```

3. LE t1 to t3 Image J 
```{r}
#mixed linear model 3
LM_LEtreat3 <- lmer(gr_t1t3 ~ treatment + (1|geno) + (1|tank), data = le)
summary(LM_LEtreat3)
anova(LM_LEtreat3)
```

### Scatter plots with regression 


#### Total calcification
1. BW from t0 to t3 w linear regression
```{r}
#define custom labels for genotype
labels_genotype <- c("C" = "Coopers",
                     "M" = "Marker 9",
                     "S" = "Sunny Isles")

#bw scatter plot w linear regression
bwgenoplot1 <-  ggplot(aes(x = ta, y = gr_t0t3, colour = geno), data = bw) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  geom_smooth(color = 1, method = 'lm', se = FALSE, linetype="dashed") +
  #scale_x_discrete(labels = custom_labels_x) + # Set custom labels for x-axis (treatment)
  scale_colour_discrete(labels = labels_genotype) +
  xlab('TA (umol)')+
  ylab("Growth rate (mg/g/day)")+
  labs(colour = "Genotype")
#scale_y_continuous(breaks = seq(0, 12, by = 1)) # Add more tick marks on the y-axis
print(bwgenoplot1)
```
2. BW from t0 to t3 with non-linear regression

Check fit of log regression... Not a great R-squared? Should I do an OLS on my linear model first and check residuals? 
```{r fit of log regression}
#check fit of log and non function
non_log.bw2 = lm(gr_t0t3 ~ta, data = bw)
summary(non_log.bw2)
plot(non_log.bw2)

log.bw2 = lm(gr_t0t3 ~ log(ta), data = bw)
summary(log.bw2)
plot(log.bw2)

#compare log to nonlog model.. 
anova(non_log.bw2,log.bw2)
```
 

```{r}
#bw scatter plot w non-linear regression (log)
bwgenoplot2 <-  ggplot(aes(x = ta, y = gr_t0t3, colour = geno), data = bw) +
  geom_point() +
  geom_smooth(method = "lm", formula = y~log(x), se = FALSE) +
  geom_smooth(color = 1, method = 'lm', formula = y~log(x), se = FALSE, linetype = "dashed") +
  scale_colour_discrete(labels = labels_genotype) +
  xlab('TA (umol)')+
  ylab("Growth rate (mg/g/day)")+
  labs(colour = "Genotype")

print(bwgenoplot2)
```
another way: 
```{r}
#need to estimate starting values
#stat_smooth(method = 'nls', formula = 'y~a*exp(b*x)',
#                method.args = list(start=c(a=0.1646, b=9.5e-8)), se=FALSE) +
```


3. BW from t1 to t3 w linear regression
```{r}
#bw scatter plot w linear regression
bwgenoplot3 <-  ggplot(aes(x = ta, y = gr_t1t3, colour = geno), data = bw) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  geom_smooth(color = 1, method = 'lm', se = FALSE, linetype="dashed") +
  #scale_x_discrete(labels = custom_labels_x) + # Set custom labels for x-axis (treatment)
  scale_colour_discrete(labels = labels_genotype) +
  xlab('TA (umol)')+
  ylab("Growth rate (mg/g/day)")+
  labs(colour = "Genotype")
#scale_y_continuous(breaks = seq(0, 12, by = 1)) # Add more tick marks on the y-axis

print(bwgenoplot3)
```

4. BW from t1 to t3 with non-linear regression 

coming soon... 

#### Linear Extension 
1. Linear extension (via 3d scanning) with linear regression
```{r}
#LE scatter plot w linear regression using Leios data
leplot1 <- ggplot(aes(x = ta, y = gr_leios, colour = geno), data = le)+
  geom_point() +
  geom_smooth(method = 'lm', se = FALSE) +
  geom_smooth(color = 1, method = 'lm', se = FALSE, linetype="dashed") +
  #scale_x_discrete(labels = custom_labels_x) + # Set custom labels for x-axis (treatment)
  scale_colour_discrete(labels = labels_genotype) +
  xlab('TA (umol)')+
  ylab("Growth rate (mm/day)")+
  labs(colour = "Genotype")

print(leplot1)
```

2. Linear extension (via 3d scanning) with non-linear regression

coming soon...

3. Linear extension (via ImageJ) from t0 to t3 with linear regression
```{r}
#LE scatter plot w linear regression Image J
leplot3 <- ggplot(aes(x = ta, y = gr_t0t3, colour = geno), data = le)+
  geom_point() +
  geom_smooth(method = 'lm', se = FALSE) +
  geom_smooth(color = 1, method = 'lm', se = FALSE, linetype="dashed") +
  #scale_x_discrete(labels = custom_labels_x) + # Set custom labels for x-axis (treatment)
  scale_colour_discrete(labels = labels_genotype) +
  xlab('TA (umol)')+
  ylab("Growth rate (mm/day)")+
  labs(colour = "Genotype")

print(leplot3)
```

4. Linear extension (via ImageJ) from t0 to t3 with non-linear regression 

coming soon... 

5. Linear extension (via ImageJ) from t1 to t3 with linear regression 

```{r}
#LE scatter plot w linear regression ImageJ
leplot5 <- ggplot(aes(x = ta, y = gr_t1t3, colour = geno), data = le)+
  geom_point() +
  geom_smooth(method = 'lm', se = FALSE) +
  geom_smooth(color = 1, method = 'lm', se = FALSE, linetype="dashed") +
  #scale_x_discrete(labels = custom_labels_x) + # Set custom labels for x-axis (treatment)
  scale_colour_discrete(labels = labels_genotype) +
  xlab('TA (umol)')+
  ylab("Growth rate (mm/day)")+
  labs(colour = "Genotype")

print(leplot5)
```

6. Linear extension (via ImageJ) from t1 to t3 with non-linear regression 

Coming soon... 
