---
title: "ALE_Experiment"
author: "Kenzie Cooke"
date: "2024-06-16"
output:
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```



```{r libraries}
library(ggplot2)
library(dplyr)
library(nlstools)
library(seacarb)
library(emmeans)

```


```{r import, echo=FALSE}
#import growth data 
bw <- read.csv('data/ALEbw6.16.csv', stringsAsFactors = TRUE)
le <- read.csv('data/ALE_le6.16.csv', stringsAsFactors = TRUE)
```
# 2/3/24 to 3/23/24 (t0-t3)  

50 days of growth.. uneven lighting until Feb 15th.  

## Total calfification by ta  

*bw data in this version normalized to initial mass and not including density corrections

### Linear mixed effect model
to check for random effects (tank, genotype)
BW from t0 to t3, growth ~ ta, 1|geno, 1|tank

```{r}
#attempting to fix "error in initializerPtr(): function 'chlomod_factor_ldetA' not provided by package 'Matrix'

#tools::package_dependencies("Matrix", which = "LinkingTo", reverse = TRUE)[[1L]]

#utils::install.packages("lme4", type = "source")

#load packages
library(lmerTest)

#BW mixed effect linear model 1
LM_BWTA1 <- lmerTest::lmer(gr_t0t3 ~ ta + (1|geno) + (1|tank), data = bw)
summary(LM_BWTA1)
anova(LM_BWTA1)
ranova(LM_BWTA1)
step(LM_BWTA1)
plot(LM_BWTA1)


```

Shows no significant tank or genotype effect. Suggests appropriate model is growth ~ ta. Create regular linear model and check residuals:

### Linear model, bw growth ~ ta
```{r}
LM_BWTA1b <- lm(gr_t0t3 ~ ta, data = bw)
summary(LM_BWTA1b)
anova(LM_BWTA1b)
plot(LM_BWTA1b)
```

## Linear extension by TA
### linear mixed effect models
#### LE via 3d scanning  
Initially scanned on Feb 1. Corals ramping up to TA treatments from Feb 3rd to Feb 6th. Corals euthanized on March 25th. So incorporating 54 days of growth.  


```{r}
#mixed effect linear model le determined via 3d scan
LM_LETA1 <- lmer(gr_leios ~ ta + (1|geno) + (1|tank), data = le)
summary(LM_LETA1)
anova(LM_LETA1)
ranova(LM_LETA1)
step(LM_LETA1)
```

No tank or genotype effect. model output of growth ~ 1 suggests TA had no significant effect on linear extension

#### LE via ImageJ (t0-t3)

```{r}
#mixed linear model 2
#assumes geno and tank are random effects
LM_LETA2 <- lmer(gr_t0t3 ~ ta + (1|geno) + (1|tank), data = le)
summary(LM_LETA2)
anova(LM_LETA2)
ranova(LM_LETA2)
step(LM_LETA2)

#adding "* geno" as a fixed effect that interacts with ta
#if added "ta + geno", would signify two fixed effects that dont interact
LM_LETA2b <- lmer(gr_t0t3 ~ ta * geno + (1|tank), data = le)
summary(LM_LETA2b)
anova(LM_LETA2b)
step(LM_LETA2b)

```

No tank or geno effect. No significant effect of ta using either:
gr_t0t3 ~ ta + (1|geno) + (1|tank)    or 
gr_t0t3 ~ ta * geno + (1|tank)

Type III analysis of variance model has p value of 0.0687 for ta in the ta*geno model..    

## Scatter plots with regression 
### Total calcification
#### Linear regression
BW from t0 to t3 w linear regression
```{r}
#define custom labels for genotype
labels_genotype <- c("C" = "Coopers",
                     "M" = "Marker 9",
                     "S" = "Sunny Isles")

#bw scatter plot w linear regression
bwgenoplot1 <-  ggplot(aes(x = ta, y = gr_t0t3, colour = geno), data = bw) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  geom_smooth(color = 1, method = 'lm', se = FALSE, linetype="dashed") +
  #scale_x_discrete(labels = custom_labels_x) + 
  #Set custom labels for x-axis (treatment)
  scale_colour_discrete(labels = labels_genotype) +
  xlab('TA (umol)')+
  ylab("Growth rate (mg/g/day)")+
  labs(colour = "Genotype")
#scale_y_continuous(breaks = seq(0, 12, by = 1)) # Add more tick marks on the y-axis
print(bwgenoplot1)
```

#### Non-linear regression

BW from t0 to t3 with non-linear regression  

Check fit of log regression:
```{r}
#check fit of log and non function
log.bw2 = lm(gr_t0t3 ~ log(ta), data = bw)
summary(log.bw2)
plot(log.bw2)

```

check fit of non log regression:

```{r}
non_log.bw2 = lm(gr_t0t3 ~ta, data = bw)
summary(non_log.bw2)
plot(non_log.bw2)
```

Compare the two models: 

```{r}
#compare log to nonlog model.. 
anova(non_log.bw2,log.bw2)
```

Similar R-squared, seems linear regression a bit better. 

Example of log regression plot: 
```{r}
#bw scatter plot w non-linear regression (log)
bwgenoplot2 <-  ggplot(aes(x = ta, y = gr_t0t3, colour = geno), data = bw) +
  geom_point() +
  geom_smooth(method = "lm", formula = y~log(x), se = FALSE) +
  geom_smooth(color = 1, method = 'lm', formula = y~log(x), se = FALSE, linetype = "dashed") +
  scale_colour_discrete(labels = labels_genotype) +
  xlab('TA (umol)')+
  ylab("Growth rate (mg/g/day)")+
  labs(colour = "Genotype")

print(bwgenoplot2)
```

coming soon... investigate some other regression fits... 
```{r}
#need to estimate starting values
#stat_smooth(method = 'nls', formula = 'y~a*exp(b*x)',
#                method.args = list(start=c(a=0.1646, b=9.5e-8)), se=FALSE) +
```



```{r}
##BW from t1 to t3 w linear regression

# bwgenoplot3 <-  ggplot(aes(x = ta, y = gr_t1t3, colour = geno), data = bw) +
#   geom_point() +
#   geom_smooth(method = "lm", se = FALSE) +
#   geom_smooth(color = 1, method = 'lm', se = FALSE, linetype="dashed") +
#   #scale_x_discrete(labels = custom_labels_x) + # Set custom labels for x-axis (treatment)
#   scale_colour_discrete(labels = labels_genotype) +
#   xlab('TA (umol)')+
#   ylab("Growth rate (mg/g/day)")+
#   labs(colour = "Genotype")
##scale_y_continuous(breaks = seq(0, 12, by = 1)) # Add more tick marks on the y-axis

#print(bwgenoplot3)
```

### Linear Extension 
#### Linear regression

Linear extension (via 3d scanning) with linear regression
```{r}
#LE scatter plot w linear regression using Leios data
leplot1 <- ggplot(aes(x = ta, y = gr_leios, colour = geno), data = le)+
  geom_point() +
  geom_smooth(method = 'lm', se = FALSE) +
  geom_smooth(color = 1, method = 'lm', se = FALSE, linetype="dashed") +
  #scale_x_discrete(labels = custom_labels_x) + # Set custom labels for x-axis (treatment)
  scale_colour_discrete(labels = labels_genotype) +
  xlab('TA (umol)')+
  ylab("Growth rate (mm/day)")+
  labs(colour = "Genotype")

print(leplot1)
```

Linear extension (via ImageJ) with linear regression (t0-t3)
```{r}
#LE scatter plot w linear regression Image J
leplot3 <- ggplot(aes(x = ta, y = gr_t0t3, colour = geno), data = le)+
  geom_point() +
  geom_smooth(method = 'lm', se = FALSE) +
  geom_smooth(color = 1, method = 'lm', se = FALSE, linetype="dashed") +
  #scale_x_discrete(labels = custom_labels_x) + # Set custom labels for x-axis (treatment)
  scale_colour_discrete(labels = labels_genotype) +
  xlab('TA (umol)')+
  ylab("Growth rate (mm/day)")+
  labs(colour = "Genotype")

print(leplot3)
```

Linear extension boxplot using IMageJ by treatment
```{r}
lebox <- boxplot(le$gr_t0t3 ~ le$treatment,
        ylab = "mm/day", xlab = "treatment") 
print(lebox)
```

# 2/19 to 3/23  (t1-t3)
After fixing lights, took new data point. 33 total days of growth  

## Total calfification by ta  

*bw data in this version normalized to initial mass and not including density corrections  

### Linear mixed effect model  

to check for random effects (tank, genotype)  
BW from t1 to t3, growth ~ ta, 1|geno, 1|tank  

```{r}
#attempting to fix "error in initializerPtr(): function 'chlomod_factor_ldetA' not provided by package 'Matrix'

#tools::package_dependencies("Matrix", which = "LinkingTo", reverse = TRUE)[[1L]]

#utils::install.packages("lme4", type = "source")

#load packages
library(lmerTest)

#BW mixed effect linear model 1
LM_BWTA33 <- lmerTest::lmer(gr_t1t3 ~ ta + (1|geno) + (1|tank), data = bw)
summary(LM_BWTA33)
anova(LM_BWTA33)
ranova(LM_BWTA33)
step(LM_BWTA33)
plot(LM_BWTA33)


```
  
Shows no significant tank or genotype effect. Suggests appropriate model is growth ~ ta. Create regular linear model and check residuals:  

### Linear model, bw growth ~ ta

```{r}
LM_BWTA33b <- lm(gr_t1t3 ~ ta, data = bw)
summary(LM_BWTA33b)
anova(LM_BWTA33b)
plot(LM_BWTA33b)
```

## Linear extension by TA 

### linear mixed effect models  

#### LE via ImageJ (t1-t3)  

```{r}
#mixed linear model 2
#assumes geno and tank are random effects
LM_LETA33 <- lmer(gr_t1t3 ~ ta + (1|geno) + (1|tank), data = le)
summary(LM_LETA33)
anova(LM_LETA33)
ranova(LM_LETA33)
step(LM_LETA33)

#adding "* geno" as a fixed effect that interacts with ta
#if added "ta + geno", would signify two fixed effects that dont interact
LM_LETA33b <- lmer(gr_t1t3 ~ ta * geno + (1|tank), data = le)
summary(LM_LETA33b)
anova(LM_LETA33b)
step(LM_LETA33b)

```

Significant effect of ta using both:  

gr_t1t3 ~ ta + (1|geno) + (1|tank)  

anova results: p=0.034  

step results: p= 0.04

gr_t1t3 ~ ta * geno + (1|tank)  

anova results: p = 0.022  

step results: p = 0.041    

## Scatter plots with regression 

### Total calcification

#### Linear regression

BW from t1 to t3 w linear regression  
```{r}
#define custom labels for genotype
labels_genotype <- c("C" = "Coopers",
                     "M" = "Marker 9",
                     "S" = "Sunny Isles")

#bw scatter plot w linear regression
bwgenoplot33 <-  ggplot(aes(x = ta, y = gr_t1t3, colour = geno), data = bw) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  geom_smooth(color = 1, method = 'lm', se = FALSE, linetype="dashed") +
  #scale_x_discrete(labels = custom_labels_x) + # Set custom labels for x-axis (treatment)
  scale_colour_discrete(labels = labels_genotype) +
  xlab('TA (umol)')+
  ylab("Growth rate (mg/g/day)")+
  labs(colour = "Genotype")
#scale_y_continuous(breaks = seq(0, 12, by = 1)) # Add more tick marks on the y-axis
print(bwgenoplot33)
```

#### Non-linear regression 

BW from t1 to t3 with non-linear regression  


Check fit of log regression:  

```{r}
#check fit of log and non function
log.bw33 = lm(gr_t1t3 ~ log(ta), data = bw)
summary(log.bw33)
plot(log.bw33)

```

check fit of non log regression  

```{r}
non_log.bw33 = lm(gr_t1t3 ~ta, data = bw)
summary(non_log.bw33)
plot(non_log.bw33)
```

Compare the two models:  

```{r}
#compare log to nonlog model.. 
anova(non_log.bw33,log.bw33)
```

outcome???  


*investigate some other regression fits... coming soon...*
```{r}
#need to estimate starting values
#stat_smooth(method = 'nls', formula = 'y~a*exp(b*x)',
#                method.args = list(start=c(a=0.1646, b=9.5e-8)), se=FALSE) +
```



```{r}
##BW from t1 to t3 w linear regression

# bwgenoplot3 <-  ggplot(aes(x = ta, y = gr_t1t3, colour = geno), data = bw) +
#   geom_point() +
#   geom_smooth(method = "lm", se = FALSE) +
#   geom_smooth(color = 1, method = 'lm', se = FALSE, linetype="dashed") +
#   #scale_x_discrete(labels = custom_labels_x) + # Set custom labels for x-axis (treatment)
#   scale_colour_discrete(labels = labels_genotype) +
#   xlab('TA (umol)')+
#   ylab("Growth rate (mg/g/day)")+
#   labs(colour = "Genotype")
##scale_y_continuous(breaks = seq(0, 12, by = 1)) # Add more tick marks on the y-axis

#print(bwgenoplot3)
```

### Linear Extension 

#### Linear regression 

Linear extension (via ImageJ) with linear regression (t1-t3)  

```{r}
#LE scatter plot w linear regression Image J
leplot33c <- ggplot(aes(x = ta, y = gr_t1t3, colour = geno), data = le)+
  geom_point() +
  geom_smooth(method = 'lm', se = FALSE) +
  geom_smooth(color = 1, method = 'lm', se = FALSE, linetype="dashed") +
  #scale_x_discrete(labels = custom_labels_x) + # Set custom labels for x-axis (treatment)
  scale_colour_discrete(labels = labels_genotype) +
  xlab('TA (umol)')+
  ylab("Growth rate (mm/day)")+
  labs(colour = "Genotype")

print(leplot33c)
```

Linear extension boxplot using IMageJ by treatment  

```{r}
lebox <- boxplot(le$gr_t1t3 ~ le$treatment,
        ylab = "mm/day", xlab = "treatment") 
print(lebox)
```

# Carb Chemistry 
```{r}
#import data
bottle_amb <-  read.csv('data/ambient_r.csv', stringsAsFactors = TRUE)

# Run seacarb - ambient, omit blanks
seacarb.output.amb = carb(
  #flag refers to which two CO2 system variables are being used; here we are using ALK and DIC
  flag = 15,
 
  #need to convert TA from umol/kg to mol/kg
  var1 = bottle_amb$calc_avg_ta / 10 ^ 6,
 
  #need to convert DIC from umol/kg to mol/kg
  var2 = bottle_amb$dic / 10 ^ 6,
 
  #salinity, in psu
  S = bottle_amb$sal,
 
  #temperature, in deg C
  T = bottle_amb$temp,
 
  #pressure, need to convert from dbar to bar
  P = bottle_amb$dbar / 10
)

#Elevated

bottle_ele <- read.csv('data/elevated_r.csv', stringsAsFactors = TRUE)
  
seacarb.output.ele = carb(
  #flag refers to which two CO2 system variables are being used; here we are using ALK and DIC
  flag = 15,
 
  #need to convert TA from umol/kg to mol/kg
  var1 = bottle_ele$calc_avg_ta / 10 ^ 6,
 
  #need to convert DIC from umol/kg to mol/kg
  var2 = bottle_ele$dic / 10 ^ 6,
 
  #salinity, in psu
  S = bottle_ele$sal,
 
  #temperature, in deg C
  T = bottle_ele$temp,
 
  #pressure, need to convert from dbar to bar
  P = bottle_ele$dbar / 10
)


#High

bottle_hi <- read.csv('data/high_r.csv', stringsAsFactors = TRUE)
  
seacarb.output.hi = carb(
  #flag refers to which two CO2 system variables are being used; here we are using ALK and DIC
  flag = 15,
 
  #need to convert TA from umol/kg to mol/kg
  var1 = bottle_hi$calc_avg_ta / 10 ^ 6,
 
  #need to convert DIC from umol/kg to mol/kg
  var2 = bottle_hi$dic / 10 ^ 6,
 
  #salinity, in psu
  S = bottle_hi$sal,
 
  #temperature, in deg C
  T = bottle_hi$temp,
 
  #pressure, need to convert from dbar to bar
  P = bottle_hi$dbar / 10
)

#Extra High

bottle_xhi <- read.csv('data/xhigh_r.csv', stringsAsFactors = TRUE)
  
seacarb.output.xhi = carb(
  #flag refers to which two CO2 system variables are being used; here we are using ALK and DIC
  flag = 15,
 
  #need to convert TA from umol/kg to mol/kg
  var1 = bottle_xhi$calc_avg_ta / 10 ^ 6,
 
  #need to convert DIC from umol/kg to mol/kg
  var2 = bottle_xhi$dic / 10 ^ 6,
 
  #salinity, in psu
  S = bottle_xhi$sal,
 
  #temperature, in deg C
  T = bottle_xhi$temp,
 
  #pressure, need to convert from dbar to bar
  P = bottle_xhi$dbar / 10
)
```

### Growth rates over time

```{r}
bwgr.time <- read.csv('data/bw_time2.csv', stringsAsFactors = TRUE)

# Convert the Date column to Date type
bwgr.time$date <- bwgr.time$date <- as.Date(bwgr.time$Date, format="%m/%d/%Y")

# Ensure the dataframe is ordered by Date
bwgr.time <- bwgr.time[order(bwgr.time$Date), ]

# Plot the data
bw_timeplot <-ggplot(bwgr.time, aes(x=date, y=B1)) +
  #geom_line(aes(color=beaker))+
  #geom_point(aes(color=beaker)) +
  xlab ("Date") +
  ylab("B1")

bw_timeplot
```

