---
title: "ALE_Experiment"
author: "Kenzie Cooke"
date: "2024-06-16"
output:
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```



```{r libraries}
library(ggplot2)
library(dplyr)
library(nlstools)
library(seacarb)
library(emmeans)
library(broom)
library(lmerTest)

```


```{r import, echo=FALSE}
#import growth data 
bw <- read.csv('data/ALEbw6.16.csv', stringsAsFactors = TRUE)
le <- read.csv('data/le_allTA.csv', stringsAsFactors = TRUE)

le.2 <- read.csv('data/le_editTA.csv', stringsAsFactors = TRUE)
#changed all negative growth to zero in le files

#bw data with corrections for density. surface area and initial mass normalized rates
bw.2 <- read.csv('data/bw_t1t3._7.20.csv', stringsAsFactors = TRUE)

bw.2$ta_mmol<-bw$ta/1000
bw.2$ta_mmol_dev<-bw$ta_dev/1000

bw$ta_mmol<-bw$ta/1000 # Create new variable. I converted umol to mmol since the slopes were 0.00 something when the [ta] were in the order of thousands 


# change the TA units for better slope magnitude 
le$ta_mmol<-le$ta/1000
le.2$ta_mmol <-le.2$ta/1000
```

# 2/3/24 to 3/23/24 (t0-t3)  

50 days of growth.. uneven lighting until Feb 15th.  

## Total calfification by ta  

*bw data in this version normalized to initial mass and not including density corrections

### Linear mixed effect model
to check for random effects (tank, genotype)
BW from t0 to t3, growth ~ ta, 1|geno, 1|tank

```{r}
#attempting to fix "error in initializerPtr(): function 'chlomod_factor_ldetA' not provided by package 'Matrix'

#tools::package_dependencies("Matrix", which = "LinkingTo", reverse = TRUE)[[1L]]

#utils::install.packages("lme4", type = "source")


#BW mixed effect linear model 1
LM_BWTA1 <- lmerTest::lmer(gr_t0t3 ~ ta_mmol + (1|geno) + (1|tank), data = bw)
summary(LM_BWTA1)
anova(LM_BWTA1)
ranova(LM_BWTA1)
step(LM_BWTA1)
plot(LM_BWTA1)

```

Shows no significant tank or genotype effect. Suggests appropriate model is growth ~ ta. Create regular linear model and check residuals:

### Linear model, bw growth ~ ta
```{r}
LM_BWTA1b <- lm(gr_t0t3 ~ ta_mmol, data = bw)
summary(LM_BWTA1b)
anova(LM_BWTA1b)
plot(LM_BWTA1b)

```

## Linear extension by TA

### linear mixed effect models

#### LE via 3d scanning  
Initially scanned on Feb 1. Corals ramping up to TA treatments from Feb 3rd to Feb 6th. Corals euthanized on March 25th. So incorporating 54 days of growth.  


```{r}
#mixed effect linear model le determined via 3d scan
LM_LETA1 <- lmer(gr_leios ~ ta_mmol + (1|geno) + (1|tank), data = le)
summary(LM_LETA1)
anova(LM_LETA1)
ranova(LM_LETA1)
step(LM_LETA1)
```

No tank or genotype effect. model output of growth ~ 1 suggests TA had no significant effect on linear extension

#### LE via ImageJ (t0-t3)

```{r}
#mixed linear model 2
#assumes geno and tank are random effects
LM_LETA2 <- lmer(gr_t0t3 ~ ta_mmol + (1|geno) + (1|tank), data = le)
summary(LM_LETA2)
anova(LM_LETA2)
ranova(LM_LETA2)
step(LM_LETA2)

#adding "* geno" as a fixed effect that interacts with ta
#if added "ta + geno", would signify two fixed effects that dont interact

LM_LETA2b <- lmer(gr_t0t3 ~ ta_mmol * geno + (1|tank), data = le)
summary(LM_LETA2b)
anova(LM_LETA2b)
step(LM_LETA2b)

```

No tank or geno effect. No significant effect of ta using either:
gr_t0t3 ~ ta + (1|geno) + (1|tank)    or 
gr_t0t3 ~ ta * geno + (1|tank)

Type III analysis of variance model has p value of 0.0687 for ta in the ta*geno model..    

## Scatter plots with regression 

### Total calcification

#### Linear regression

BW from t0 to t3 w linear regression
```{r}
#define custom labels for genotype
labels_genotype <- c("C" = "Coopers",
                     "M" = "Marker 9",
                     "S" = "Sunny Isles")

#bw scatter plot w linear regression
bwgenoplot1 <-  ggplot(aes(x = ta, y = gr_t0t3, colour = geno), data = bw) +
  geom_point(aes(shape=treatment)) +
  geom_smooth(method = "lm", se = FALSE) +
  geom_smooth(color = 1, method = 'lm', se = FALSE, linetype="dashed") +
  #scale_x_discrete(labels = custom_labels_x) + 
  #Set custom labels for x-axis (treatment)
  scale_colour_discrete(labels = labels_genotype) +
  xlab('TA (umol)')+
  ylab("Growth rate (mg/g/day)")+
  labs(colour = "Genotype")
#scale_y_continuous(breaks = seq(0, 12, by = 1)) # Add more tick marks on the y-axis


# KENZIE, print is not needed for a ggplot plot. Just calling the name works 
bwgenoplot1
```


```{r}
#linear regression value for each genotype
TA_gr <- bw %>%
                group_by(geno) %>%
                do({model = lm(gr_t0t3~ta_mmol, data=.)  # model
                    data.frame(tidy(model), # get coefficient info
                    glance(model))})        # get model info
TA_gr

# While run individually, Coopers slope is marginally significant (p=0.053)
```


#### Non-linear regression

BW from t0 to t3 with non-linear regression  

Check fit of log regression:
```{r}
#check fit of log and non function
log.bw2 = lm(gr_t0t3 ~ log(ta_mmol), data = bw)
summary(log.bw2)
plot(log.bw2)

```

check fit of non log regression:

```{r}
non_log.bw2 = lm(gr_t0t3 ~ta_mmol, data = bw)
summary(non_log.bw2)
plot(non_log.bw2)
```

Compare the two models: 

```{r}
#compare log to nonlog model.. 
anova(non_log.bw2,log.bw2)
```

Similar R-squared, seems linear regression a bit better. 

Example of log regression plot: 
```{r}
#bw scatter plot w non-linear regression (log)
bwgenoplot2 <-  ggplot(aes(x = ta, y = gr_t0t3, colour = geno), data = bw) +
  geom_point(aes(shape=treatment)) +
  geom_smooth(method = "lm", formula = y~log(x), se = FALSE) +
  geom_smooth(color = 1, method = 'lm', formula = y~log(x), se = FALSE, linetype = "dashed") +
  scale_colour_discrete(labels = labels_genotype) +
  xlab(expression(TA~(mu~mol)))+ # This is ANA, Changed this for better mu notation
  ylab(expression(Growth~rate~(mg~g^-1~day^-1)))+ # Changed this for better notation
  labs(colour = "Genotype")

bwgenoplot2
```

coming soon... investigate some other regression fits... 
```{r}
#need to estimate starting values
#stat_smooth(method = 'nls', formula = 'y~a*exp(b*x)',
#                method.args = list(start=c(a=0.1646, b=9.5e-8)), se=FALSE) +
```


### Linear Extension 

#### Linear regression

Linear extension (via 3d scanning) with linear regression
```{r}
#LE scatter plot w linear regression using Leios data
leplot1 <- ggplot(aes(x = ta, y = gr_leios, colour = geno), data = le)+
  geom_point(aes(shape = treatment)) +
  geom_smooth(method = 'lm', se = FALSE) +
  geom_smooth(color = 1, method = 'lm', se = FALSE, linetype="dashed") +
  #scale_x_discrete(labels = custom_labels_x) + # Set custom labels for x-axis (treatment)
  scale_colour_discrete(labels = labels_genotype) +
  xlab(expression(TA~(mu~mol))) +
  ylab(expression(Growth~rate~(mm~day^-1)))+
  labs(colour = "Genotype")

print(leplot1)
```

Linear extension (via ImageJ) with linear regression (t0-t3)
```{r}
#LE scatter plot w linear regression Image J
leplot3 <- ggplot(aes(x = ta, y = gr_t0t3, colour = geno), data = le)+
  geom_point(aes(shape = treatment)) +
  geom_smooth(method = 'lm', se = FALSE) +
  geom_smooth(color = 1, method = 'lm', se = FALSE, linetype="dashed") +
  #scale_x_discrete(labels = custom_labels_x) + # Set custom labels for x-axis (treatment)
  scale_colour_discrete(labels = labels_genotype) +
  xlab(expression(TA~(mu~mol))) +
  ylab(expression(Growh~rate~(mm~day^-1)))+
  labs(colour = "Genotype")
  
print(leplot3)
```

Linear extension boxplot using IMageJ by treatment
```{r}
lebox <- boxplot(le$gr_t0t3 ~ le$treatment,
        ylab = expression(mm~day^-1), xlab = "Treatment") 
print(lebox)
```

# 2/19 to 3/23  (t1-t3)
After fixing lights, took new data point. 33 total days of growth  

## Total calfification by ta  
Wet weight converted to dry weight, corrected for density and standardized to initial surface area.

### Linear mixed effect model  

to check for random effects (tank, genotype)  
BW from t1 to t3, growth ~ ta, 1|geno, 1|tank  

```{r}
#attempting to fix "error in initializerPtr(): function 'chlomod_factor_ldetA' not provided by package 'Matrix'

#tools::package_dependencies("Matrix", which = "LinkingTo", reverse = TRUE)[[1L]]

#utils::install.packages("lme4", type = "source")

#load packages
library(lmerTest)

#with mean ta derived from 2/20-end, p becomes 0.056.......!!!!!!!

#BW mixed effect linear model 1
LM_BWTA33 <- lmerTest::lmer(gr_sa ~ ta_mmol + (1|geno) + (1|tank), data = bw.2)
summary(LM_BWTA33)
anova(LM_BWTA33)
ranova(LM_BWTA33)
step(LM_BWTA33)
plot(LM_BWTA33)

```
  
Shows no significant tank or genotype effect. Suggests appropriate model is growth ~ ta. Create regular linear model and check residuals:  

### Linear model, bw growth ~ ta

```{r}
LM_BWTA33b <- lm(gr_sa ~ ta_mmol, data = bw.2)
summary(LM_BWTA33b)
anova(LM_BWTA33b)
plot(LM_BWTA33b)
```

## Linear extension by TA 

### linear mixed effect models  

#### LE via ImageJ (t1-t3)  

```{r}
#mixed linear model 2


#assumes geno and tank are random effects
LM_LETA33 <- lmer(gr_t1t3 ~ ta_mmol + (1|geno) + (1|tank), data = le)
summary(LM_LETA33)
anova(LM_LETA33)
ranova(LM_LETA33)
step(LM_LETA33)

#adding "* geno" as a fixed effect that interacts with ta
#if added "ta + geno", would signify two fixed effects that dont interact

LM_LETA33b <- lmer(gr_t1t3 ~ ta_mmol * geno + (1|tank), data = le)
summary(LM_LETA33b)
anova(LM_LETA33b)
step(LM_LETA33b)
```
Significant effect of ta on linear extension, no sig effect of tank or geno. 

Most simple model: gr_t1t3 ~ ta
anova results: p=0.032  
```{r}

# Most simple model 
LM_LETA33c <- lm(gr_t1t3 ~ ta_mmol, data = le)
summary(LM_LETA33c)
anova(LM_LETA33c)



```

Individual regressions by genotype (S, C, M):

```{r}

# Adding this for individual regressions

TA_LE <- le %>%
          group_by(geno) %>%
                do({model = lm(gr_t1t3~ta_mmol, data=.)  # model
                    data.frame(tidy(model), # get coefficient info
                    glance(model))})        # get model info
TA_LE

# this is Ana The individual regressions are not significant. I guess that does not matter  since genotype was not significant, so in theory all data should be ran together?
```
Ana: The individual regressions are not significant. I guess that does not matter since genotype was not significant, so in theory all data should be ran together?  


## Scatter plots with regression 

### Total calcification

#### Linear regression

BW from t1 to t3 w linear regression  
Growth standardized to initial surface area: 
```{r}
#define custom labels for genotype
labels_genotype <- c("C" = "Coopers",
                     "M" = "Marker 9",
                     "S" = "Sunny Isles")

#bw scatter plot w linear regression
bwgenoplot33 <-  ggplot(aes(x = ta, y = gr_sa, colour = geno), data = bw.2) +
  geom_point(aes(shape = treatment)) +
  geom_smooth(method = "lm", se = FALSE) +
  geom_smooth(color = 1, method = 'lm', se = FALSE, linetype="dashed") +
  #scale_x_discrete(labels = custom_labels_x) + # Set custom labels for x-axis (treatment)
  scale_colour_discrete(labels = labels_genotype) +
   xlab(expression(TA~(mu~mol)))+ 
  ylab(expression(Growth~rate~(mg~cm^2 ^-1~day^-1)))+ 
  labs(colour = "Genotype")+
  theme_classic() # Removes grey background and change the axes... there are so many themes that you can try, but I like this becuase is is clean and simple. theme_bw() is good too

bwgenoplot33
```
  

Growth standardized to initial mass:  

```{r}
#bw scatter plot w linear regression
bwgenoplot33mass <-  ggplot(aes(x = ta, y = gr_mass, colour = geno), data = bw.2) +
  geom_point(aes(shape = treatment)) +
  geom_smooth(method = "lm", se = FALSE) +
  geom_smooth(color = 1, method = 'lm', se = FALSE, linetype="dashed") +
  #scale_x_discrete(labels = custom_labels_x) + # Set custom labels for x-axis (treatment)
  scale_colour_discrete(labels = labels_genotype) +
   xlab(expression(TA~(mu~mol)))+ 
  ylab(expression(Growth~rate~(mg~g^-1~day^-1)))+ 
  labs(colour = "Genotype")+
  theme_classic() # Removes grey background and change the axes... there are so many themes that you can try, but I like this becuase is is clean and simple. theme_bw() is good too
bwgenoplot33mass
```

#### Non-linear regression 

BW from t1 to t3 with non-linear regression    
growth standardized to surface area

Check fit of log regression:  

```{r}
#check fit of log and non function
log.bw33 = lm(gr_sa ~ log(ta), data = bw.2)
summary(log.bw33)
plot(log.bw33)
```

check fit of non log regression  

```{r}
non_log.bw33 = lm(gr_sa ~ta, data = bw.2)
summary(non_log.bw33)
plot(non_log.bw33)
```

Compare the two models:  

```{r}
#compare log to nonlog model.. 
anova(non_log.bw33, log.bw33)
```

outcome???  

Similar R-squared, seems linear regression a bit better. 

Example of log regression plot: 
```{r}

#bw scatter plot w non-linear regression (log)
bwgenoplotlog <-  ggplot(aes(x = ta, y = gr_sa, colour = geno), data = bw.2) +
  geom_point(aes(shape=treatment)) +
  geom_smooth(method = "lm", formula = y~log(x), se = FALSE) +
  geom_smooth(color = 1, method = 'lm', formula = y~log(x), se = FALSE, linetype = "dashed") +
  scale_colour_discrete(labels = labels_genotype) +
  xlab(expression(TA~(mu~mol)))+ # This is ANA, Changed this for better mu notation
  ylab(expression(Growth~rate~(mg~cm^2 ^-1~day^-1)))+ # Changed this for better notation
  labs(colour = "Genotype") +
  theme_classic()

bwgenoplotlog
```


### Linear Extension 

#### Linear regression 

Linear extension (via ImageJ) with linear regression (t1-t3)  

```{r}
#LE scatter plot w linear regression Image J
leplot33c <- ggplot(aes(x = ta, y = gr_t1t3, colour = geno), data = le)+
  geom_point(aes(shape = treatment)) +
  geom_smooth(method = 'lm', se = FALSE) +
  geom_smooth(color = 1, method = 'lm', se = FALSE, linetype="dashed") +
  scale_colour_discrete(labels = labels_genotype) +
  xlab(expression(TA~(mu~mol)))+ 
  ylab(expression(Growth~rate~(mm~day^-1)))+ 
  labs(colour = "Genotype")+
  theme_classic()

leplot33c
```

Adding error bars for TA:  
```{r}
#LE scatter plot w linear regression Image J
leplot33cc <- ggplot(aes(x = ta, y = gr_t1t3, colour = geno), data = le.2)+
  geom_point(aes(shape = treatment)) +
  geom_errorbarh(aes(xmin = ta - ta_se, xmax = ta + ta_)) +
  geom_smooth(method = 'lm', se = FALSE) +
  geom_smooth(color = 1, method = 'lm', se = FALSE, linetype="dashed") +
  scale_colour_discrete(labels = labels_genotype) +
  xlab(expression(TA~(mu~mol)))+ 
  ylab(expression(Growth~rate~(mm~day^-1)))+ 
  labs(colour = "Genotype")+
  theme_classic()

leplot33cc
```





# Carb Chemistry  

## Statistics

Was alkalinity statistically different between treatments?   

```{r}

treat_aov <- aov(ta_mmol ~ treatment, data = bw.2)
summary(treat_aov)

tukey_treat <- TukeyHSD(treat_aov)
print(tukey_treat)

```
yes...  

## bar graph 

Bar graph of treatment TA

```{r}
ta_time <- read.csv('data/ta_time.csv', stringsAsFactors = TRUE)

library(dplyr)
library(ggplot2)

ta_summary <- ta_time %>%
  group_by(treatment) %>%
  summarise(
    mean_ta = mean(ta_umol),
    sem_ta = sd(ta_umol) / sqrt(n())
  )
print(ta_summary)

box_ta <- ggplot(ta_summary, aes(x = treatment, y = mean_ta)) +
  geom_bar(stat = "identity", position = position_dodge(), width = 0.7) +
  geom_errorbar(aes(ymin = mean_ta - sem_ta, ymax = mean_ta + sem_ta),width = 0.2,position = position_dodge(0.7)) +
  theme_classic()

box_ta

```


## Stability over time  

Stability of treatment over time:
```{r}


#convert date column to "Date" type.
ta_time$date <- as.Date(ta_time$date, format = "%m/%d/%Y")

labels_treatment <- c("ambient" = "ambient",
                     "elevated" = "elevated",
                     "high" = "high",
                     "xhigh" = "extra high")

custom_colors <- c("ambient" = "#999999", "elevated" = "#00AFBB", "high" = "#E7B800", "xhigh" = "#FC4E07")

# Reverse the order of the treatments in the legend
treatment_order <- rev(names(custom_colors))

# Get the unique dates from the data
unique_dates <- unique(ta_time$date)

treat_stab <- ggplot(ta_time, aes(x = date, y = ta_umol, colour = treatment)) +
  geom_point() +
  scale_colour_manual(values = custom_colors, limits=treatment_order) +
  scale_x_date(breaks = unique_dates, date_labels = "%m/%d/%Y") +
  ylab(expression(TA ~ (mu ~ mol))) +  
  theme_classic()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


treat_stab
```


Stability of treatments over time with mean and error bars: 

```{r}
# Summarize the data to create mean and std dev for each day by treatment

summary_data <- ta_time %>%
  group_by(date, treatment) %>%
  summarise(
    mean_ta = mean(ta_umol),
    se_ta = sd(ta_umol) / sqrt(n())
  ) %>%
  ungroup()

# Plot the summarized data
treat_stab2 <- ggplot(summary_data, aes(x = date, y = mean_ta, colour = treatment)) +
  geom_point(size=3) +
  geom_linerange(aes(ymin = mean_ta - se_ta, ymax = mean_ta + se_ta)) +
  scale_colour_manual(values = custom_colors, limits=treatment_order) +
    scale_x_date(breaks = unique_dates, date_labels = "%m/%d/%Y") +
  ylab(expression(TA ~ (mu ~ mol))) +  
  theme_classic()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


treat_stab2
```


## Blanks
Total alkalinity in blanks by treatment  
```{r}
bl <- read.csv('data/blanks.csv', stringsAsFactors = TRUE)

blanks_time <- ggplot(subset(bl, treatment %in% "AMB", color = beaker), aes(date, ta_avg)) +
  geom_point() +

blanks_time
```



```{r}
#import data
bottle_amb <-  read.csv('data/ambient_r.csv', stringsAsFactors = TRUE)

# Run seacarb - ambient, omit blanks
seacarb.output.amb = carb(
  #flag refers to which two CO2 system variables are being used; here we are using ALK and DIC
  flag = 15,
 
  #need to convert TA from umol/kg to mol/kg
  var1 = bottle_amb$calc_avg_ta / 10 ^ 6,
 
  #need to convert DIC from umol/kg to mol/kg
  var2 = bottle_amb$dic / 10 ^ 6,
 
  #salinity, in psu
  S = bottle_amb$sal,
 
  #temperature, in deg C
  T = bottle_amb$temp,
 
  #pressure, need to convert from dbar to bar
  P = bottle_amb$dbar / 10
)


#Elevated

bottle_ele <- read.csv('data/elevated_r.csv', stringsAsFactors = TRUE)
  
seacarb.output.ele = carb(
  #flag refers to which two CO2 system variables are being used; here we are using ALK and DIC
  flag = 15,
 
  #need to convert TA from umol/kg to mol/kg
  var1 = bottle_ele$calc_avg_ta / 10 ^ 6,
 
  #need to convert DIC from umol/kg to mol/kg
  var2 = bottle_ele$dic / 10 ^ 6,
 
  #salinity, in psu
  S = bottle_ele$sal,
 
  #temperature, in deg C
  T = bottle_ele$temp,
 
  #pressure, need to convert from dbar to bar
  P = bottle_ele$dbar / 10
)


#High

bottle_hi <- read.csv('data/high_r.csv', stringsAsFactors = TRUE)
  
seacarb.output.hi = carb(
  #flag refers to which two CO2 system variables are being used; here we are using ALK and DIC
  flag = 15,
 
  #need to convert TA from umol/kg to mol/kg
  var1 = bottle_hi$calc_avg_ta / 10 ^ 6,
 
  #need to convert DIC from umol/kg to mol/kg
  var2 = bottle_hi$dic / 10 ^ 6,
 
  #salinity, in psu
  S = bottle_hi$sal,
 
  #temperature, in deg C
  T = bottle_hi$temp,
 
  #pressure, need to convert from dbar to bar
  P = bottle_hi$dbar / 10
)

#Extra High

bottle_xhi <- read.csv('data/xhigh_r.csv', stringsAsFactors = TRUE)
  
seacarb.output.xhi = carb(
  #flag refers to which two CO2 system variables are being used; here we are using ALK and DIC
  flag = 15,
 
  #need to convert TA from umol/kg to mol/kg
  var1 = bottle_xhi$calc_avg_ta / 10 ^ 6,
 
  #need to convert DIC from umol/kg to mol/kg
  var2 = bottle_xhi$dic / 10 ^ 6,
 
  #salinity, in psu
  S = bottle_xhi$sal,
 
  #temperature, in deg C
  T = bottle_xhi$temp,
 
  #pressure, need to convert from dbar to bar
  P = bottle_xhi$dbar / 10
)
```

